\chapter{EC2 lab}

\section{VPC creation}

\begin{enumerate}

\item 
  Create the VPC named \texttt{LAB\_VPC} with IP range 10.0.0.0/16.

\item
  Create the Subnet named \texttt{LAB\_SUBNET\_1} with:
[6~  \begin{description}
  \item[IP range] 10.0.1.0/24
  \item[automatic public IP assignment] turned on
  \end{description}
  
\item
  Create the Internet Gateway named \texttt{LAB\_GATEWAY}.

\item
  Attach the Internet Gateway to the VPC.

\item
  Find the route table for the VPC.
  Add a route to send traffic for anywhere (0.0.0.0/0) to your internet gateway.

\item
  Create the Security Group named \texttt{LAB\_SG} with description \texttt{LAB\_SG}.

\item 
  Add a rule to your security group permit SSH traffic (TCP port 22) inbound from anywhere on the internet (0.0.0.0/0).
  
\end{enumerate}

\section{SSH key setup}

\begin{enumerate}
\item Open your public key using PuttyGen.
\item Copy the public key (for pasting into OpenSSH authorized keys)
\item Go to EC2 and then Key Pairs in the AWS Console.
\item Choose Import Key
\item Name the key pair MAIN\_KEY. 
\item Paste the key you copied into the box.
\end{enumerate}

\section{EC2 instance}

Create EC2 instance as follows:
\begin{enumerate}
\item Open the instances page by browsing to Services, EC2, instances.

\item Hit Launch Instances.

\item Name the instance \texttt{LAB\_INSTANCE}.

\item Pick Amazon Linux from the AMI list.

\item Check that instance type is \texttt{t2.micro} (should be default)

\item Under Key Pair drop down to \texttt{MAIN\_KEY}.

\item On Network Settings hit Edit. Choose your \texttt{LAB\_VPC} from the list.

\item Ensure that your \texttt{LAB\_SUBNET\_1} is selected.

\item Ensure that Auto-assign public IP is set to Enable.
  
\item Choose Select existing security group.

\item From the list pick \texttt{LAB\_SG}.

\item Hit Launch Instance in the bottom right.

\item Click the Instances link and notice that your instance will be Pending before changing to Running.
  
\end{enumerate}

\section{SSH to your instance}

\begin{enumerate}
\item Click into your instance in the list.

\item Click the copy button beside Public IPv4 address.

\item Open PuTTY, type the EC2 IP address into the box.

\item Go to connection, data and put \texttt{ec2-user} under the auto login username.

\item Go to connection, SSH, Auth and choose the private key file in the private key file for authentication box. 

\item
The first time you connect to a host you'll get a warning:

\begin{verbatim}
The authenticity of host '54.78.220.233 (54.78.220.233)' can't be established.
ECDSA key fingerprint is SHA256:8omkD5RLibZNgJJ/B7MAnL7IbEcrmCmIWFdQXbjJf60.
Are you sure you want to continue connecting (yes/no)?
\end{verbatim}

This is OK!

\item 
If you see something like the following then you're connected:

\begin{verbatim}
       __|  __|_  )
       _|  (     /   Amazon Linux 2 AMI
      ___|\___|___|
\end{verbatim}

\end{enumerate}

\section{Cleanup}

Delete all resources you made in this lab:
\begin{enumerate}
\item Terminate the instance.
\item Delete the VPC.
\end{enumerate}

